On First-order Proofs without syntax

### 2021-01-03 main.tex
Fix the macros

\newcommand{\Gr}{\mathcal{G}}
\newcommand{\LG}{\mathcal{L}}
\newcommand{\FT}{\mathcal{F}}
\newcommand{\US}{\mathcal{U}}

@@

\begin{definition}
The propositional encoding $A^P$ of a formula $A$ is defined inductively by:

\begin{centering}
	$a^P = a$ for every atom $a$

	$(A \vee B)^P = A^P \vee B^P$ \hspace{2cm} $(A \wedge B)^P = A^P \wedge B^P$

	$(\forall x A)^P = U_x \vee A^P$ \hspace{2cm} $(\exists x A)^P = E_x \wedge A^P$

@@

Similarly, we can define the propositional encoding $S^P$ of a context $S$ inductively by setting $\Box^P = \Box$. Note that $S^P$ is also an encoding.

@@

\begin{proposition}
For any context $S$ and any formula $A$:
\begin{itemize}
  \item $A^P$ is a formula containing no quantifier for any formula $A$.
  \item $\Gr(A^P) = \Gr(A)$ by confounding the atoms $U_x$, $E_x$ with the variable
	  $x$. Thus, a map $f : \Gr(A^P) \rightarrow \Gr(B^P)$ can be seen as a map
		$f : \Gr(A) \rightarrow \Gr(B)$.
  \item $(S\{A\})^P = S^P\{A^P\}$.
\end{itemize}

@@

\begin{proposition}
\label{prop311}
Let $A$ and $B$ be two formulas such that
$\od{\odd{\odh {A}}{}{B}{\{\wD, \cD\}}}$. Then 
$\od{\odd{\odh {A^P}}{}{B^P}{\{\wD, \cD\}}}$.
\end{proposition}

\iffalse
\begin{proof}

We say that a rule \begin{prooftree} \hypo{C} \infer1[]{D} \end{prooftree}
is $P$-derivable for $\{\wD, \cD\}$ if
        $\od{\odd{\odh {C^P}}{}{D^P}{\{\wD, \cD\}}}$.

We now prove that $\wD$ and $\cD$ are
	$P$-derivable for $\{\wD, \cD\}$.

\begin{itemize}
\item $\wD$ is $P$-derivable: we have
  \begin{prooftree}
    \hypo{S^P\{\fff\}}
    \infer1[$\wD$]{S^P\{C^P\}}
  \end{prooftree}

\item $\cD$ is $P$-derivable: we have
  \begin{prooftree}
    \hypo{S^P\{C^P \vee C^P\}}
    \infer1[$\cD$]{S^P\{C^P\}}
  \end{prooftree}

Hence, $\od{\odd{\odh {A^P}}{}{B^P}{\{\wD, \cD\}}}$.

\end{proof}
\fi

@@

Now we prove that each map $g \in \{wk, ac, m, m_1, m_2 \}$ is a skew bifibration. To do that, it suffices to prove that $g$ is a fibration between the corresponding binding graphs since it is already a skew fibration on the corresponding fographs and it is label-preserving and existential-preserving.
\begin{center}
for each $x$-binder $b$ in $\Gr(B^P)$, for each vertex $v \in V(\Gr(A^P))$
such that $g(v)$ is bound by $b$, there exists a unique binder $b'$ such
that $b'$ binds $v$.
\end{center}

@@

$f$ can be seen as a skew fibration from $G(A^P)$ to $G(B^P)$, which gives the existence of the propositions $A'$ and $B'$, and of the following derivation:
  \[\od{\odd{\odd{\odd{\odh{A^P} }
  {\Delta }{A'}{\me} }
  {\Delta' }{B'}{\acD} }
  {\Delta''}{B^P}{\wD}} \]

\begin{lemma} there exists $B''$ such that $B''^P$ = $B'$.

\begin{proof}
Consider the dervation $\Delta''$. If some $U_x$ (or $E_x$) is introduced
via weakening, then all the atoms it binds in $B^P$ should also be introduced 
via weakening. In fact, an atom of $B^P$ is introduced via weakening is 
equivalent to the fact that its corresponding vertex is not in the image of $f$. 
Since there is an edge from $U_x$ (resp. $E_x$) to all the literals it binds in the 
binding graph $\overrightarrow{G(B)}$, if one of the atoms is in the image, 
$U_x$ (resp. $E_x$) should also be in the image since $f$ is a fibration on binding graphs.

@@

Then the derivation:
 \[\od{\odd{\odd{\odh{A^P} }
  {\Delta }{A'}{\me} }
  {\Delta' }{B'}{\acD}} \]
can be translated to the derivation:
\[\od{\odd{\odh{(A^P)^*} }
{\Delta^* }{(B')^*}{}} \]

@@

$\Delta^*$ can now be transformed into a valid derivation $\Delta_1$ by using the two
transformation rules above and by applying them in a bottom-up style:
\[\od{\odd{\odh{(A^P)^*} }
{\Delta_1 }{(B')^*}{\acD, \me, \me'}} \]

@@

Suppose that $(C \vee E) \wedge (D \vee F) = G^P$ for some $G$.
Since $C \vee E$ cannot be $Q_x$ (otherwise, the rule applied would be
$\me'$), $G$ can be written as $G_1 \wedge G_2$ with $C \vee E = G_1^P$ and $D
\vee F = G_2^P$.

@@

If $G_i = \forall x H_i$ for some $i$, then there will be a conjunction of $U_x^*$
and some formula which can never be eliminated by the rules $\me$, $\me'$ and
$\acD$. However, there exists no such conjunction in ${A^P}^*$, which leads to a
contradiction.

Hence, $G_i$ can be written as $J_i \vee K_i$ for $i = 1, 2$. We now have $(C
\wedge D) \vee (E \wedge F) = ((J_1 \wedge J_2) \vee (K_1 \wedge K_2))^P$.

@@

Suppose that $Q_X^* \wedge (C \vee D)$ = $F^P$ for some $P$. It is clear
that $Q_x^* = E_x^*$ and $F = \exists x G$ with $G^P = C \vee D$ for some $G$.
We distinguish two cases:
\begin{itemize}
  \item $G = \forall y H$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
D)$ has a subformula $(E_x^* \wedge U_y^*)$, which cannot be eliminated by the
rules $\me$, $\me'$, $\acD$. It is clear that ${A^P}^*$ does not
have a subformula of this form, which leads to a contradiction.
  \item $G = G_1 \vee G_2$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
	  D) = ((\exists x G_1) \vee (\exists x G_2))^P$.
\end{itemize}

>>>>

\newcommand{\Gr}{\mathcal{G}}
\newcommand{\LG}{\mathcal{L}}
\newcommand{\FT}{\mathcal{F}}
\newcommand{\US}{\mathcal{U}}
\newcommand{\PE}[1]{#1^P}

@@

\begin{definition}
The propositional encoding $\PE{A}$ of a formula $A$ is defined inductively by:

\begin{centering}
	$\PE{a} = a$ for every atom $a$

	$\PE{(A \vee B)} = \PE{A} \vee \PE{B}$ \hspace{2cm} $\PE{(A \wedge B)} =
	\PE{A} \wedge \PE{B}$

	$\PE{(\forall x A)} = U_x \vee \PE{A}$ \hspace{2cm} $\PE{(\exists x A)}
	= E_x \wedge \PE{A}$

@@

Similarly, we can define the propositional encoding $\PE{S}$ of a context $S$
inductively by setting $\PE{\Box} = \Box$. Note that $\PE{S}$ is also a context.

@@

\begin{proposition}
For any context $S$ and any formula $A$:
\begin{itemize}
  \item $\PE{A}$ is a formula containing no quantifier for any formula $A$.
  \item $\Gr(\PE{A}) = \Gr(A)$ by confounding the atoms $U_x$, $E_x$ with the variable
	  $x$. Thus, a map $f : \Gr(\PE{A}) \rightarrow \Gr(\PE{B})$ can be seen as a map
		$f : \Gr(A) \rightarrow \Gr(B)$.
  \item $\PE{(S\{A\})} = \PE{S}\{\PE{A}\}$.
\end{itemize}

@@

\begin{proposition}
\label{prop311}
Let $A$ and $B$ be two formulas such that
$\od{\odd{\odh {A}}{}{B}{\{\wD, \cD\}}}$. Then 
$\od{\odd{\odh {\PE{A}}}{}{\PE{B}}{\{\wD, \cD\}}}$.
\end{proposition}

@@

Now we prove that each map $g \in \{wk, ac, m, m_1, m_2 \}$ is a skew bifibration. To do that, it suffices to prove that $g$ is a fibration between the corresponding binding graphs since it is already a skew fibration on the corresponding fographs and it is label-preserving and existential-preserving.
\begin{center}
for each $x$-binder $b$ in $\Gr(\PE{B})$, for each vertex $v \in V(\Gr(\PE{A}))$
such that $g(v)$ is bound by $b$, there exists a unique binder $b'$ such
that $b'$ binds $v$.
\end{center}

@@

$f$ can be seen as a skew fibration from $G(\PE{A})$ to $G(\PE{B})$, which gives the existence of the propositions $A'$ and $B'$, and of the following derivation:
  \[\od{\odd{\odd{\odd{\odh{\PE{A}} }
  {\Delta }{A'}{\me} }
  {\Delta' }{B'}{\acD} }
  {\Delta''}{\PE{B}}{\wD}} \]

\begin{lemma} there exists $B''$ such that $\PE{B''}$ = $B'$.

\begin{proof}
Consider the dervation $\Delta''$. If some $U_x$ (or $E_x$) is introduced
via weakening, then all the atoms it binds in $\PE{B}$ should also be introduced 
via weakening. In fact, an atom of $\PE{B}$ is introduced via weakening is 
equivalent to the fact that its corresponding vertex is not in the image of $f$. 
Since there is an edge from $U_x$ (resp. $E_x$) to all the literals it binds in the 
binding graph $\overrightarrow{G(B)}$, if one of the atoms is in the image, 
$U_x$ (resp. $E_x$) should also be in the image since $f$ is a fibration on binding graphs.

@@

Then the derivation:
 \[\od{\odd{\odd{\odh{\PE{A}} }
  {\Delta }{A'}{\me} }
  {\Delta' }{B'}{\acD}} \]
can be translated to the derivation:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta^* }{(B')^*}{}} \]

@@

$\Delta^*$ can now be transformed into a valid derivation $\Delta_1$ by using the two
transformation rules above and by applying them in a bottom-up style:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta_1 }{(B')^*}{\acD, \me, \me'}} \]

@@

Suppose that $(C \vee E) \wedge (D \vee F) = \PE{G}$ for some $G$.
Since $C \vee E$ cannot be $Q_x$ (otherwise, the rule applied would be
$\me'$), $G$ can be written as $G_1 \wedge G_2$ with $C \vee E = \PE{G_1}$ and $D
\vee F = \PE{G_2}$.

@@

If $G_i = \forall x H_i$ for some $i$, then there will be a conjunction of $U_x^*$
and some formula which can never be eliminated by the rules $\me$, $\me'$ and
$\acD$. However, there exists no such conjunction in ${\PE{A}}^*$, which leads to a
contradiction.

Hence, $G_i$ can be written as $J_i \vee K_i$ for $i = 1, 2$. We now have $(C
\wedge D) \vee (E \wedge F) = \PE{((J_1 \wedge J_2) \vee (K_1 \wedge
K_2))}$.

@@

Suppose that $Q_X^* \wedge (C \vee D)$ = $\PE{F}$ for some $F$. It is clear
that $Q_x^* = E_x^*$ and $F = \exists x G$ with $\PE{G} = C \vee D$ for some $G$.
We distinguish two cases:
\begin{itemize}
  \item $G = \forall y H$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
D)$ has a subformula $(E_x^* \wedge U_y^*)$, which cannot be eliminated by the
	rules $\me$, $\me'$, $\acD$. It is clear that ${\PE{A}}^*$ does not
have a subformula of this form, which leads to a contradiction.
  \item $G = G_1 \vee G_2$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
  D) = \PE{((\exists x G_1) \vee (\exists x G_2))}$.
\end{itemize}
